---
title: "543 assignment 3"
output:
  html_document:
    df_print: paged
---

# Data Source
This assignment would use the same data as the assignment 1 and assignment 2.
  
  
#### 1. Data processing  
#### 1.1 Import the map  

```{r}
linkMap="https://github.com/EvansDataScience/VAforPM_Spatial/raw/main/worldMap.geojson" 

library(sf)
mapWorld=read_sf(linkMap)

# see data in map
head(mapWorld)
# the projection: WGS 84
```

#### 1.2 Show the Base Layer
```{r}
library(ggplot2)
# plot original map
base=ggplot(data=mapWorld) + geom_sf(fill='grey90',
                                     color=NA) + theme_classic()

base #base map
```
  
#### 1.3 Import the Data
```{r}
link = 'https://raw.githubusercontent.com/tianyl27/543_coursework_2/main/WHR2021.csv'

# getting the data TABLE from the file:
WHRData2021 <- read.csv(file=url(link))

# Rename some columns:
colnames(WHRData2021)[1]='NAME'
colnames(WHRData2021)[8]='GDP'
colnames(WHRData2021)[10]='LifeExpectancy'
colnames(WHRData2021)[23]='COVIDDeathRate'
colnames(WHRData2021)[31]='GINI'

head(WHRData2021)
```
  
#### 2. Cluster
#### 2.1 Boxplot of interested variables  

```{r}

boxplot(WHRData2021[,c('GDP','LifeExpectancy','COVIDDeathRate','GINI')])

```

#### 2.2 Radar plot  

```{r}
library(reshape2)

WHRData2021_sub = WHRData2021[,c('NAME','GDP','LifeExpectancy','COVIDDeathRate','GINI')]

# change the wide format to the long format
WHRData2021_long=melt(WHRData2021_sub,id.vars = 'NAME')
head(WHRData2021_long)

base2  = ggplot(WHRData2021_long, 
               aes(x = variable, 
                   y = value, 
                   group = NAME)) + #new
        geom_polygon(fill = 'gray',
                     col='orange') 

plot1 = base2 + coord_polar()

plot2 = plot1 + facet_wrap(~NAME,# one plot per city
                           ncol = 20) # 20 plot per row
#plot2
```

#### 2.3 Standardize the variables  

```{r}

# standardize: all variable share the same range of value
WHRData2021$GDP_S=as.vector(scale(WHRData2021$GDP))
WHRData2021$LE_S=as.vector(scale(WHRData2021$LifeExpectancy))
WHRData2021$CDR_S=as.vector(scale(WHRData2021$COVIDDeathRate))


set.seed(121)

library(cluster)
vars=c('GDP_S','LE_S','CDR_S')#variables needed

distMatrix=cluster::daisy(WHRData2021[,vars])
          
res.pam=cluster::pam(x=distMatrix,
                     k = 4,
                     cluster.only = F)


WHRData2021$cluster=as.factor(res.pam$clustering)

theVars=c('GDP_S','LE_S','CDR_S','cluster')
aggregate(.~cluster,
          data=WHRData2021[,theVars],
          FUN=median)

aggregate(.~cluster,
          data=WHRData2021[,theVars],
          FUN=mean)

```

#### 2.4 Rename the clusters  
Rename each cluster according to the median and mean results.  

```{r}
WHRData2021$cluster=factor(WHRData2021$cluster,
                           levels=c(3,1,2,4),
                           labels=c("Developed and less deaths","Developed but most deaths","Moderate","Developing but least deaths"), 
                           ordered=T)
```


#### 3 Mapping

#### 3.1 Merge the Data
Merge the data with the map data.  

```{r}
mapWorldVars=merge(mapWorld, #map first
                   WHRData2021, 
                   by='NAME') 

```

#### 3.2 Mapping Univariate Numerical
Use the Ladder score, which is the comprehensive index of this report, to draw a choropleth map.  

```{r}
colMap= base + geom_sf(data=mapWorldVars,
                       aes(fill=Ladder.score),
                       color=NA)
colMap
```

#### 3.3 Mapping the clustering result  

```{r}
TitleText = "Country Performance in 2020 COVID epidemic"
CaptionText = "Source: https://worldhappiness.report/ed/2021/"
theLegTitle="World_Performance\n(grey is missing)"

clusterMap= base + geom_sf(data=mapWorldVars,
                       aes(fill=cluster),
                       color=NA)

clusterMap+ scale_fill_brewer(palette ='BuGn',
                              direction = -1,
                              name=theLegTitle)+ labs(title=TitleText,
                                                      caption=CaptionText) +
  theme(plot.title = element_text(hjust = 1), 
        plot.caption = element_text(hjust = 0))

```

